An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-29 draft.tex
Some small corrections and adds acks

The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of  Function Evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.

@@

\begin{table}
    \small
    \caption{10 Populations experiment parameters
    }
    \label{tab:params:10}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{5 Populations experiment parameters
    }
    \label{tab:params:5}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

@@

>>>>

The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.

@@

\begin{table}
    \small
    \caption{Parameters used in the experiments with ten populations
    }
    \label{tab:params:10}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{Parameters used in the experiments that used five populations
    }
    \label{tab:params:5}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

@@

\section{Acknowledgements}

This paper has been supported in part by projects DeepBio
(TIN2017-85727-C4-2-P).
